Supplemental Material: Multidimensional analysis and detection of informative features in human brain white matter

### 2021-04-20 reports/manuscript/paper_changes.tex
ENH: Add responses to reviewer comments

\usepackage{changes}

@@

colby2012, ODonnell2009-uu}. While this retains important information
about each individual's white matter, it also presents statistical
challenges due to the dimensionality of the data. In past work,
comparisons between groups or across individuals were done
independently at each node of each bundle, for each

@@

%\end{enumerate}
The present work aims to balance predictive accuracy with descriptive power
\cite{Murdoch2019-ax, Breiman2001-uz} by capitalizing on all of the available
data across all bundles, while also retaining and elucidating spatial
information about the locations that are most informative for discriminative
performance. To do so, we use a linear modeling approach, which aims to
redict phenotypical variance in a group of subjects (or classify group
membership), based on a linear combination of the features estimated with
tractometry.

@@

Using this approach, we first need to deal with the large and
asymmetric dimensionality of the data: tractometry data usually has
many more features (i.e., number of measurements per individual) than
samples (number of subjects), which makes inferences from the data
about phenotypical differences between individuals ill-posed. This
regime is the target of several statistical learning techniques, and
is often solved by various forms of regularization.

% For example, Tikhonov regularization shrinks the solution such that the sum of
% squared contributions from the individual features are minimized
% \cite{Hoerl2000-ij}. Another solution to the problem is provided by
% the
The Lasso algorithm minimizes the sum of the absolute values of
contributions of each feature \cite{Tibshirani1996-qs}. This
tends to shrink to zero the contributions of many of the features,
providing results that are both accurate and interpretable. When
additional structure is available in the organization of the data,
regularization algorithms can take advantage of this structure. For
example, if the features lend themselves to a natural division into
different groups, the Group Lasso (GL) can be used to select groups
of features, rather than individual features \cite{Yuan2006-ky}.
The Sparse Group Lasso (SGL) elaborates on this idea by providing
control both of group sparsity, as well as overall sparsity of the
solutions \cite{simon2013sgl}. Because the features measured with
tractomery lend themselves to grouping based on the tracts from which
each measurement is taken, GL and SGL provide useful tools
for linear model fitting in problems of this form. Here we, first,
develop an implementation of SGL that is well suited to the analysis of
tractometry data and, second, demonstrate the power and flexibility of
this approach by applying it to both classification
and continuous prediction problems. Our tractometry data flow is
represented in \cref{fig:methods} and explained in further detail in
the methods section.

\begin{figure}[h!]
    \vspace{0.25cm}
    \rlap{
        \hspace{0.25em}
        \includegraphics[width=0.98\columnwidth]{methods-brains.pdf}
    }
    \vspace{-8.25cm}
    \includegraphics[width=\columnwidth]{method-quad.pdf}
    {\phantomsubcaption\label{fig:methods:tractogram}}
    {\phantomsubcaption\label{fig:methods:cst}}
    {\phantomsubcaption\label{fig:methods:tract-profile}}
    {\phantomsubcaption\label{fig:methods:group-structure}}
    \caption{{\bf Tractometry data flow}
        \label{fig:methods}
        {\bf (a)} Whole brain tractography generates streamlines approximating
        the trajectories of white matter connections.
        {\bf (b)} Tractometry classifies these streamlines into anatomical bundles.
        In this case, we show the left corticospinal tract (CSTL) over a mid-saggital
        anatomical slice.
        {\bf (c)} Tractometry further extracts bundle profiles,
        quantifications of various diffusion metrics along the length of the
        fiber bundle. Here, we show one subject's fractional anisotropy (FA)
        profile for CSTL.
        {\bf (d)} the phenotypical target data and tractometric features can
        be organized into a linear model, $\hat{y} = \mathbf{X}
        \hat{\beta}$. The feature matrix $\mathbf{X}$ is color-coded
        to reveal a natural group structure: the left (orange) group
        contains $k$ features from the CSTL, the middle (green) group
        contains $k$ features from the left cingulum cingulate (CGCL),
        and the right (blue) group
        contains $k$ features from the left arcuate (ARCL). The coefficients in
        $\hat{\beta}$ follow the same natural grouping.
    }
\end{figure}

@@

\begin{figure}
    \vspace{0.0cm}
    \rlap{
        \includegraphics[width=0.33\textwidth]{als_coefs.pdf}
    }
    \vspace{-7.0cm}
    \includegraphics[width=\textwidth]{als_results.pdf}
    {\phantomsubcaption\label{fig:class-results:probs}}
    {\phantomsubcaption\label{fig:class-results:coefs3d}}
    {\phantomsubcaption\label{fig:class-results:tract-profiles}}
    {\phantomsubcaption\label{fig:class-results:false-positives}}
    {\phantomsubcaption\label{fig:class-results:false-negatives}}
    \caption{%
        {\bf SGL accurately and interpretably predicts ALS diagnosis.}
        \label{fig:class-results}
        {\bf (a)} Classification probabilities for ALS diagnosis, with
        controls on the left, patients on the right, predicted controls in
        blue, and predicted patients in orange. That is, orange dots on the left
        represent false positives, while blue dots on the right represent
        false negatives. We achieve {\protect\alsAccuracy}\% accuracy with an
        ROC AUC of {\protect\alsRocAuc}.
        {\bf (b)} SGL coefficients are presented on the core fibers of major
        fiber bundles. They exhibit high group sparsity and are concentrated
        in the FA of the CST. The brain is oriented with the right hemisphere
        in the foreground and anterior to the right of the page. The CSTL,
        CSTR, and CFA bundles are indicated for orientation.
        {\bf (c)} SGL identifies three portions of the CST as important,
        where $\hat{\beta}$ (dashed line, right axis) has large values. These
        are centered around nodes 30, 65, and 90, corresponding to locations
        of substantial differences in FA between the ALS and control groups
        (shaded areas indicates standard error of the mean).
        {\bf (d)} Bundle profiles for false positive classifications. Line
        colors correspond to the marker edge color in the top left plot.
        These individuals have reduced FA in the CST portions which SGL
        identified as important. Their misclassification is coherent with the
        feature importance and the group differences in FA.
        {\bf (e)} Individual bundle profiles for false negative
        classifications. These individuals have bundle profiles which
        oscillate between the group means.
    }
\end{figure}

We developed a method for analyzing dMRI tractometry data that uses the Sparse
Group Lasso (SGL) to select features that are sparse both at the group (bundle) level, as well as overall.
We demonstrate the use of this method on four different datasets in
both a classification setting and a regression setting.

@@

Using data from a previous study of patients with amyotrophic lateral sclerosis (ALS)~\cite{sarica2017corticospinal}, we tested the performance of SGL in a
classification setting. The previous study predicted ALS status with a mean
accuracy of 80\% using a random forest algorithm based on \emph{a priori}
selection of features only within the CST bundle-of-interest.
SGL delivers improved predictive performance, with a cross-validated accuracy
of {\alsAccuracy}\% and an area under the receiver operating characteristic
curve (ROC AUC) of {\alsRocAuc}, without the need for \emph{a priori} feature
engineering (\cref{fig:class-results:probs}). In addition to this
classification performance, SGL also identifies the white matter tracts most
important for ALS classification. The relative importance of white matter
features is captured in the $\beta$ coefficients from \cref{eq:sgl}.
\Cref{fig:class-results:tract-profiles} depicts these coefficients along the
right CST, plotted over the FA values for the control and ALS subject groups.
We find that SGL selects FA metrics in the corticospinal tract and
particularly in the right corticospinal tract as most important to ALS
classification, confirming previous findings \cite{van2011upper,
toosy2003diffusion, sarica2014tractography, sage2007quantitative,
sage2009quantitative, karlsborg2004corticospinal, ellis1999diffusion,
cosottini2005diffusion, ciccarelli2009investigation, abe2010voxel} and
identifying the portions of the brain that were selected \emph{a priori} in
the previous study from which we obtained the data
\cite{sarica2017corticospinal}.

To assess the added value of incorporating anatomical knowledge into our
model, we compared our classification results to those obtained from
\replaced{four other models: a pure lasso model, an elastic net model, a bundle-mean lasso (i.e. a lasso model trained only on the mean metric values from each bundle), and a principal components regression lasso (Lasso PCR) model \cite{powell2018local, rasero2021integrating}}{a lasso model}.
\replaced{These models achieved accuracies of 76\%, 76\%, 79.5\%, and 71.5\% respectively and ROC AUC values of 0.76, 0.75, 0.82, and 0.71, respectively}{The pure lasso model achieved 76\% accuracy and an ROC AUC of 0.76}
using the same cross-validation strategy used to assess the SGL model.
This difference in performance justifies the additional complexity of the SGL
over simpler sparsity-inducing regression strategies.

The $\beta$ coefficients exhibit high bundle level sparsity; only some
bundles are important, which can be be confirmed by observing the value of
$\alpha$, the regularization hyperparameter that controls the mixture of the
Group Lasso and lasso penalties, selected through nested cross-validation
(see \cref{eq:sgl}).
If $\alpha$ is closer to zero, it indicates that the phenotype in question
preferentially correlates with only a few groups of covariates. For the ALS
dataset, $\alpha = $ \alsLRatio, confirming that the white matter correlates
of ALS reside mostly in one bundle, namely the CST.

@@

To test the performance of SGL with tractometry data in a continuous
regression task, we focus here on the prediction of biological age in three
datasets named WH, HBN, and Cam-CAN (see the methods section for a description of each dataset). Prediction of ``brain
age'' is a commonly undertaken task in neuroimaging machine learning, in part
because these predictions, and deviations therefrom, may be diagnostic of
overall brain health (for a review, see Cole et al.~\cite{Cole2019-rz}).
However, as Nelson et al.~\cite{nelson2019biomarkers} have observed, aging
biomarkers are subject to unique challenges and tend to be noncausative. Our
interest in aging here relies on its utility as a methodological benchmark.
Biological age operates on a natural scale, with meaningful and easily
understood units, and it's popularity as a machine learning target makes it
valuable for comparisons with other studies.

The WH \cite{yeatman2014lifespan}, HBN \cite{alexander2017open}, and Cam-CAN
\cite{shafto2014cambridge,taylor2017cambridge} datasets used here contain
data from 76, 978, and 640 subjects, respectively, ranging from 6-50, 5-21,
and 18-88 years of age, respectively. In each case, biological age was used
as the predicted variable $y$. SGL was fit to the tractometry-extracted
features FA and mean diffusivity (MD) in 18 major brain tracts, with diffusion metrics extracted
from diffusion tensor imaging (DTI) for the WH dataset and diffusion kurtosis
imaging (DKI) \cite{jensen2005diffusion} for the HBN and Cam-CAN datasets.

@@

@@

We present here a novel method for analysis of dMRI tractometry data that
relies on the Sparse Group Lasso \cite{simon2013sparse} to make accurate
predictions of phenotypic properties of individual subjects while,
simultaneously, identifying the features of the white matter that are most
important for this prediction. The method is data-driven and it is broadly
applicable to a wide range of research questions: it performs well in
predicting both continuous variables, such as biological age, as well as
categorical variables, such as whether a person is a patient or a healthy
control. In both of these cases, SGL out-performs previous algorithms that
have been developed for these tasks \cite{sarica2017corticospinal,
Richard2018-ux, mcpherson2020single}. The nested cross-validation approach
used to fit the model and make both predictions and inferences from the model
guards against overfitting and tunes the degree of sparseness required by the
algorithm. This means that SGL can accurately describe phenomena that are
locally confined to a particular anatomical location or diffusion property
(e.g., FA in the CST) as well as phenomena that are widely distributed
amongst brain regions and measured diffusion properties.

Specifically, we demonstrated that the algorithm correctly detects the fact
that ALS, which is a disease of lower motor neurons, is localized to the
cortico-spinal tract. This recapitulates the results of previous analysis of
these same data, using a bundle-of-interest approach
\cite{sarica2017corticospinal}. In contrast, for the analysis of biological
age, the coefficients identified by the algorithm are very widely distributed
across many parts of the white matter, mirroring previous results that show a
large and continuous distribution of life-span changes in white matter
properties \cite{yeatman2014lifespan}.

@@

Taken together, these results demonstrate the promise of the
group-regularized regression approach. Even at the scale of dozens of
subjects, the results provided by SGL are both accurate, as well as
interpretable \cite{Murdoch2019-ax}: tractometry capitalizes on domain
knowledge to engineer meaningful features; SGL scores these features based on
their relative importance; enables a visualization of these feature
importance scores in the anatomical coordinate frame of the bundles (e.g.,
\cref{fig:class-results:coefs3d,fig:age-results:wh-coefs3d,fig:age-results:hbn-coefs3d,fig:age-results:cc-coefs3d})
and provides a means to understand model errors (e.g.,
\cref{fig:class-results:false-positives,fig:class-results:false-negatives}).
Thus, this multivariate analysis approach achieves high cross-validated
accuracy for precision medicine applications of dMRI data and identifies
relevant features of brain anatomy that can further our neuroscientific
understanding of clinical disorders.

Neuroscience has entered an era in which consortium efforts are amassing
large datasets of high-quality dMRI measurements to address a variety of
scientific questions \cite{jernigan2016ping, jernigan2018abcd,
alexander2017open, Miller2016-hw, VanEssen2012}, but the volume and
complexity of these data pose substantial challenges. Tractometry followed by
analysis with SGL provide a promising approach to distill meaningful insights
from the wealth of data measured in these efforts.

@@

\item Diffusion MRI data from the Cambridge Centre for Ageing and
Neuroscience (Cam-CAN) ``CC700'' dataset
\cite{shafto2014cambridge,taylor2017cambridge}, containing data from 640
subjects with ages 18-88. These data were measured on a 3T Siemens TIM Trio
system and written informed consent was obtained from each participant. Voxel
resolution was \num{2x2x2}$\text{mm}^3$ with 30 non-colinear directions
measured for each of $b=1000$ $\frac{\text{sec}}{\text{mm}^2}$ and $b=2000$
$\frac{\text{sec}}{\text{mm}^2}$. The diffusion weighted images were acquired
with a twice refocused spin-echo sequence and the same preprocessing pipeline
used for the HBN dataset was applied to this data. We will refer to this
dataset as Cam-CAN.

@@

>>>>

\lstset{basicstyle=\ttfamily}

\usepackage[commentmarkup=footnote]{changes}

@@

colby2012, ODonnell2009-uu},
\added{hereafter referred to as tract profiling}.
While this retains important information about each individual's white
matter, it also presents statistical challenges due to the dimensionality of
the data. In past work, comparisons between groups or across individuals were
done independently at each node of each bundle, for each

@@

%\end{enumerate}

\begin{figure}[b!]
    \vspace{0.25cm}
    \rlap{
        \hspace{-0.25em}
        \includegraphics[width=0.98\columnwidth]{methods-brains.pdf}
    }
    \vspace{-8.4cm}
    \includegraphics[width=\columnwidth]{method-quad.pdf}
    {\phantomsubcaption\label{fig:methods:tractogram}}
    {\phantomsubcaption\label{fig:methods:cst}}
    {\phantomsubcaption\label{fig:methods:tract-profile:cst}}
    {\phantomsubcaption\label{fig:methods:tract-profile:fa}}
    {\phantomsubcaption\label{fig:methods:group-structure}}
    \caption{\textbf{Tractometry data flow}
        \label{fig:methods}
        \textbf{(a)} Whole brain tractography generates streamlines approximating
        the trajectories of white matter connections.
        \textbf{(b)} Tractometry classifies these streamlines into anatomical bundles.
        In this case, we show the left corticospinal tract (CSTL)
        \added{and the left arcuate fasciculus (ARCL)}
        over a mid-saggital anatomical slice.
        \deleted{\textbf{(c)}} \replaced{Tract profiling}{Tractometry} further extracts bundle profiles,
        quantifications of various diffusion metrics along the length of the
        fiber bundle. Here, we show one subject's fractional anisotropy (FA)
        profile for \added{\textbf{(c)}} the CSTL \added{and \textbf{(d)} the ARCL}.
        \textbf{(d)} the phenotypical target data and \replaced{tract profile}{tractometric}
        features can be organized into a linear model, $\hat{y} = \mathbf{X}
        \hat{\beta}$. The feature matrix $\mathbf{X}$ is color-coded
        to reveal a natural group structure: the left (orange) group
        contains $k$ features from the CSTL, the middle (green) group
        contains $k$ features from the left cingulum cingulate (CGCL),
        and the right (blue) group
        contains $k$ features from the \replaced{ARCL}{left arcuate (ARCL)}.
        The coefficients in $\hat{\beta}$ follow the same natural grouping.
    }
\end{figure}

The present work aims to balance predictive accuracy with descriptive power
\cite{Murdoch2019-ax, Breiman2001-uz} by capitalizing on all of the available
data across all bundles, while also retaining and elucidating spatial
information about the locations that are most informative for discriminative
performance. To do so, we use a linear modeling approach, which aims to
redict phenotypical variance in a group of subjects (or classify group
membership), based on a linear combination of the features estimated with
\replaced{tract profiling}{tractometry}.

@@

Using this approach, we first need to deal with the large and
asymmetric dimensionality of the data: \replaced{tract profile}{tractometry}
data usually has
many more features (i.e., number of measurements per individual) than
samples (number of subjects), which makes inferences from the data
about phenotypical differences between individuals ill-posed. This
regime is the target of several statistical learning techniques, and
is often solved by various forms of regularization.

% For example, Tikhonov regularization shrinks the solution such that the sum of
% squared contributions from the individual features are minimized
% \cite{Hoerl2000-ij}. Another solution to the problem is provided by
% the
The Lasso algorithm minimizes the sum of the absolute values of
contributions of each feature \cite{Tibshirani1996-qs}. This
tends to shrink to zero the contributions of many of the features,
providing results that are both accurate and interpretable. When
additional structure is available in the organization of the data,
regularization algorithms can take advantage of this structure. For
example, if the features lend themselves to a natural division into
different groups, the Group Lasso (GL) can be used to select groups
of features, rather than individual features \cite{Yuan2006-ky}.
The Sparse Group Lasso (SGL) elaborates on this idea by providing
control both of group sparsity, as well as overall sparsity of the
solutions \cite{simon2013sgl}. Because the features measured with
tractomery lend themselves to grouping based on the tracts from which
each measurement is taken, GL and SGL provide useful tools
for linear model fitting in problems of this form. Here we, first,
develop an implementation of SGL that is well suited to the analysis of
\replaced{tract profile}{tractometry} data and, second, demonstrate the power and flexibility of
this approach by applying it to both classification
and continuous prediction problems. Our \deleted{tractometry} data flow is
represented in \cref{fig:methods} and explained in further detail in
the methods section.

\added{One more approach to multivariate analysis of neuroimaging data is to perform inference or prediction in a transformed feature space rather then the original diffusion metric feature space.}
\added{For example, in Lasso PCR, diffusion features can be projected onto a principal components (PC) basis for use in a sparsity constrained model \mbox{\cite{powell2018local, rasero2021integrating}}.}
\added{While these approaches achieve high predictive performance and may even yield biologically interpretable principal components \mbox{\cite{chamberland2019dimensionality}}, they neglect anatomical grouping information.}
\added{We view these approaches as complementary to the SGL-based approach: PCR based approaches seeks to model brain-phenotype relationships using}
\added{the most parsimonious representation of variance in the diffusion measures. In contrast, our SGL-based approach seeks to establish whether prior knowledge of anatomical grouping improves modeling of brain-phenotype relationships.}
\added{Motivated by this distinction, we also introduce a union of the two approaches called PCR-SGL, in which each group of features is independently transformed into its PC basis, thereby retaining the anatomical grouping information of}
\added{the original feature space. We demonstrate circumstances in which this approach both helps and hinders predictive performance.}

@@

We developed a method for analyzing dMRI \replaced{tract profile}{tractometry} data that uses the Sparse
Group Lasso (SGL) to select features that are sparse both at the group (bundle) level, as well as overall.
We demonstrate the use of this method on four different datasets in
both a classification setting and a regression setting.

@@

Using data from a previous study of patients with amyotrophic lateral sclerosis (ALS)~\cite{sarica2017corticospinal}, we tested the performance of SGL in a
classification setting. The previous study predicted ALS status with a mean
accuracy of 80\% using a random forest algorithm based on \emph{a priori}
selection of features only within the CST bundle-of-interest.
SGL delivers improved predictive performance, with a cross-validated accuracy
of {\alsAccuracy}\% and an area under the receiver operating characteristic
curve (ROC AUC) of {\alsRocAuc}, without the need for \emph{a priori} feature
engineering.
\added{We also predicted ALS diagnosis using the PCR-SGL, wherein each}
\added{group of features is independently transformed to its PC basis and}
\added{an SGL model is fit to the transformed features. This model achieves}
\added{\protect{\alsAccuracyGpca}\% accuracy and an ROC AUC of \protect{\alsRocAucGpca}}
(\cref{fig:class-results:probs}). In addition to this
classification performance, both SGL \added{and PCR-SGL} also \replaced{identify}{identifies}
the white matter tracts most important for ALS classification.
\added{\Cref{fig:class-results:coefs3d} shows that PCR-SGL identified as potential disease biomarkers the diffusion measures in the CST from the cerebral peduncle to the corona radiata, agreeing with the previous study from which these data were extracted \mbox{\cite{sarica2017corticospinal}}.}
The relative importance of white matter
features is captured in the $\beta$ coefficients from \cref{eq:sgl}.
\Cref{fig:class-results:tract-profiles} depicts these coefficients along the
right CST, plotted over the FA values for the control and ALS subject groups.
We find that SGL \replaced{and PCR-SGL select}{selects} FA metrics in the corticospinal tract and
particularly in the right corticospinal tract as most important to ALS
classification, confirming previous findings \cite{van2011upper,
toosy2003diffusion, sarica2014tractography, sage2007quantitative,
sage2009quantitative, karlsborg2004corticospinal, ellis1999diffusion,
cosottini2005diffusion, ciccarelli2009investigation, abe2010voxel} and
identifying the portions of the brain that were selected \emph{a priori} in
the previous study from which we obtained the data
\cite{sarica2017corticospinal}.

\begin{figure}[t!]
    \vspace{3.0cm}
    \rlap{
        \includegraphics[width=0.33\textwidth]{als_gpca_coefs.pdf}
    }
    \vspace{-7.0cm}
    \includegraphics[width=\textwidth]{als_gpca_results.pdf}
    {\phantomsubcaption\label{fig:class-results:probs}}
    {\phantomsubcaption\label{fig:class-results:coefs3d}}
    {\phantomsubcaption\label{fig:class-results:tract-profiles}}
    {\phantomsubcaption\label{fig:class-results:false-positives}}
    {\phantomsubcaption\label{fig:class-results:false-negatives}}
    \caption{%
        {\bf \added{PCR-}SGL accurately and interpretably predicts ALS diagnosis.}
        \label{fig:class-results}
        {\bf (a)} Classification probabilities for ALS diagnosis, with
        controls on the left, patients on the right, predicted controls in
        blue, and predicted patients in orange. That is, orange dots on the left
        represent false positives, while blue dots on the right represent
        false negatives. We achieve {\protect\alsAccuracy}\% accuracy with an
        ROC AUC of {\protect\alsRocAuc}.
        {\bf (b)} \added{PCR-}SGL coefficients are presented on the core fibers of major
        fiber bundles. They exhibit high group sparsity and are concentrated
        in the FA of the \replaced{corticospinal tract (CST)}{CST}.
        The brain is oriented with the right hemisphere
        in the foreground and anterior to the right of the page. The CSTL, CSTR,
        \replaced{callosum forceps anterior (CFA), left arcuate (ARCL), and right arcuate (ARCR)}{CFA}
        bundles are indicated for orientation.
        {\bf (c)} \added{PCR-}SGL identifies three portions of the CST as important,
        where $\hat{\beta}$ (dashed line, right axis) has large values. These
        are centered around nodes 30, 65, and 90, corresponding to locations
        of substantial differences in FA between the ALS and control groups
        (shaded areas indicates standard error of the mean).
        {\bf (d)} Bundle profiles for false positive classifications. Line
        colors correspond to the marker edge color in the top left plot.
        These individuals have reduced FA in the CST portions which SGL
        identified as important. Their misclassification is coherent with the
        feature importance and the group differences in FA.
        {\bf (e)} Individual bundle profiles for false negative
        classifications. These individuals have bundle profiles which
        oscillate between the group means.
    }
\end{figure}

To assess the added value of incorporating anatomical knowledge into our
model, we compared our classification results to those obtained from
\replaced{four other models: a pure lasso model, an elastic net model, a bundle-mean lasso (i.e. a lasso model trained only on the mean metric values from each bundle), and a principal components regression lasso (Lasso PCR) model \cite{powell2018local, rasero2021integrating}}{a lasso model}.
\replaced{These models achieved accuracies of 76\%, 76\%, 79.5\%, and 71.5\% respectively and ROC AUC values of 0.76, 0.75, 0.82, and 0.71, respectively}{The pure lasso model achieved 76\% accuracy and an ROC AUC of 0.76}
using the same cross-validation strategy used to assess the SGL model.
This difference in performance justifies the additional complexity of the SGL
over simpler sparsity-inducing regression strategies.
\added{The relatively poor performance of Lasso PCR suggests that features with small variance are nonetheless relevant for predicting ALS diagnosis.}
\added{Or equivalently, that group differences between ALS diagnoses are not the dominant source of variance in the diffusion metrics.}

The $\beta$ coefficients exhibit high bundle level sparsity; only some
bundles are important, which can be be confirmed by observing the value of
$\alpha$, the regularization hyperparameter that controls the mixture of the
Group Lasso and lasso penalties, selected through nested cross-validation
(see \cref{eq:sgl}).
If $\alpha$ is closer to zero, it indicates that the phenotype in question
preferentially correlates with only a few groups of covariates. For the ALS
dataset, \added{the SGL model has} $\alpha = $ \alsLRatio 
\added{and the PCR-SGL model has \protect$\alpha = $ \protect\alsLRatioGpca},
confirming that the white matter correlates of ALS reside mostly in one
bundle, namely the CST.

@@

To test the performance of SGL \deleted{with tractometry data} in a continuous
regression task, we focus here on the prediction of biological age in three
datasets named WH, HBN, and Cam-CAN (see the methods section for a description of each dataset). Prediction of ``brain
age'' is a commonly undertaken task in neuroimaging machine learning, in part
because these predictions, and deviations therefrom, may be diagnostic of
overall brain health (for a review, see Cole et al.~\cite{Cole2019-rz}).
However, as Nelson et al.~\cite{nelson2019biomarkers} have observed, aging
biomarkers are subject to unique challenges and tend to be noncausative. Our
interest in aging here relies on its utility as a methodological benchmark.
Biological age operates on a natural scale, with meaningful and easily
understood units, and it's popularity as a machine learning target makes it
valuable for comparisons with other studies.

The WH \cite{yeatman2014lifespan}, HBN \cite{alexander2017open}, and Cam-CAN
\cite{shafto2014cambridge,taylor2017cambridge} datasets used here contain
data from 76, 1651, and 640 subjects, respectively, ranging from 6-50, 5-21,
and 18-88 years of age, respectively. In each case, biological age was used
as the predicted variable $y$. SGL was fit to the \replaced{tract profile}{tractometry-extracted}
features FA and mean diffusivity (MD) in 18 major brain tracts, with diffusion metrics extracted
from diffusion tensor imaging (DTI) for the WH dataset and diffusion kurtosis
imaging (DKI) \cite{jensen2005diffusion} for the HBN and Cam-CAN datasets.

\begin{figure}[t!]
    \vspace{3.0cm}
    \rlap{
        \hspace{0.775cm}
        \includegraphics[width=0.26\textwidth]{wh_coefs.pdf}
        \hspace{0.5em}
        \includegraphics[width=0.26\textwidth]{hbn_coefs.pdf}
        \hspace{0.5em}
        \includegraphics[width=0.26\textwidth]{camcan_coefs.pdf}
    }
    \vspace{-6.1cm}
    \includegraphics[width=\textwidth]{regression_scatter.pdf}
    {\phantomsubcaption\label{fig:age-results:wh-scatter}}
    {\phantomsubcaption\label{fig:age-results:hbn-scatter}}
    {\phantomsubcaption\label{fig:age-results:cc-scatter}}
    {\phantomsubcaption\label{fig:age-results:wh-coefs3d}}
    {\phantomsubcaption\label{fig:age-results:hbn-coefs3d}}
    {\phantomsubcaption\label{fig:age-results:cc-coefs3d}}
    {\phantomsubcaption\label{fig:age-results:wh-profile}}
    {\phantomsubcaption\label{fig:age-results:hbn-profile}}
    {\phantomsubcaption\label{fig:age-results:cc-profile}}
    \caption{%
        {\bf Predicting age with tractometry and SGL.}
        \label{fig:age-results}
        {\bf (top)} The predicted age vs. true age of each individual from the test
        splits (i.e., when each subject's data was held out in fitting the
        model) for the {\bf (a)} WH, {\bf (b)} HBN, and {\bf (c)} Cam-CAN
        datasets; an accurate prediction falls close to the $y=x$ line
        (dashed). The mean absolute error (MAE) and coefficient of
        determination $R^2$ are presented in the lower right of each scatter
        plot.
        {\bf (middle)} Feature importance for predicting age from
        \replaced{tract profile}{tractometry} in
        the {\bf (d)} WH, {\bf (e)} HBN, and {\bf (f)} Cam-CAN datasets.
        The orientation of the
        brain is that same as in \cref{fig:class-results:coefs3d}, however because
        the coefficients exhibit high global sparsity (as opposed to group
        sparsity), we plot the mean of the absolute value of $\hat{\beta}$
        for each bundle on the core fiber. The global distrubution of the
        $\hat{\beta}$ coefficients reflects the fact that aging is not
        confined to a single white matter bundle.
        {\bf (bottom)} Age quintile bundle profiles for the {\bf (g)} WH,
        {\bf (h)} HBN, and {\bf (i)} Cam-CAN datasets.
    }
\end{figure}

@@

\begin{figure}[t!]
    \includegraphics[width=\textwidth]{model_performance.pdf}
    \caption{%
        {\bf \added{Model performance across all datasets.}}
        \label{fig:model-performance}
        \added{Each panel shows model performance measured on the test set for each}
        \added{cross-validation split, with each black dot representing a split, box}
        \added{plots representing the quartiles, and white diamonds representing the mean}
        \added{performance. The $y$-scale varies in each subplot.}
        \added{\textbf{(a)} Accuracy of test set predictions for the}
        \added{ALS dataset. Because group differences}
        \added{in ALS diagnosis are mostly confined to a single bundle, the group}
        \added{structure-preserving methods, SGL and PCR-SGL, outperform the other}
        \added{models.}
        \added{The remaining frames show coefficient of determination, \protect$R^2$ in test sets}
        \added{for the \textbf{(b)} WH, \textbf{(c)} HBN, and \textbf{(c)} Cam-CAN}
        \added{datasets. Because aging affects the white matter globally, group structure-blind}
        \added{methods like elastic net and PCR Lasso perform well. Nonetheless, the SGL}
        \added{models show competitve predictive performance, adapting to a problem where}
        \added{group structure is not as informative. PCR-SGL performs poorly in this regime}
        \added{because its initial group-wise PC projection destroys between bundle covariance.}
        \added{The bundle-mean lasso performs poorly, demonstrating the value of along-tract profiling.}
    }
\end{figure}
\added{Model performance across all four datasets and all six model types is summarized}
\added{in \cref{fig:model-performance}. In contrast to the ALS classification case, the PCR Lasso performs}
\added{competitively in age regression, suggesting that aging is a significant source of global variance}
\added{in the tract profiles. On the other hand, the PCR-SGL models destroy cross-bundle covariance information}
\added{in their initial group-wise PC projection step (see \cref{eq:svd}), limiting performance}
\added{for globally distributed phenomena. Conversely, the SGL models are able to adapt to this regime.}

@@

We present here a novel method for analysis of dMRI \replaced{tract profile}{tractometry} data that
relies on the Sparse Group Lasso \cite{simon2013sparse} to make accurate
predictions of phenotypic properties of individual subjects while,
simultaneously, identifying the features of the white matter that are most
important for this prediction. The method is data-driven and it is broadly
applicable to a wide range of research questions: it performs well in
predicting both continuous variables, such as biological age, as well as
categorical variables, such as whether a person is a patient or a healthy
control. In both of these cases, SGL out-performs previous algorithms that
have been developed for these tasks \cite{sarica2017corticospinal,
Richard2018-ux, mcpherson2020single}. The nested cross-validation approach
used to fit the model and make both predictions and inferences from the model
guards against overfitting and tunes the degree of sparseness required by the
algorithm. This means that SGL can accurately describe phenomena that are
locally confined to a particular anatomical location or diffusion property
(e.g., FA in the CST) as well as phenomena that are widely distributed
amongst brain regions and measured diffusion properties.

Specifically, we demonstrated that the algorithm correctly detects the fact
that ALS, which is a disease of lower motor neurons, is localized to the
cortico-spinal tract. This recapitulates the results of previous analysis of
these same data, using a bundle-of-interest approach
\cite{sarica2017corticospinal}.
\added{As with the original study, it is unclear whether our strong predictive performance based on image properties that would be visible in patient scans or whether it is identifying a new subclinical disease biomarker.} 
In contrast, for the analysis of biological
age, the coefficients identified by the algorithm are very widely distributed
across many parts of the white matter, mirroring previous results that show a
large and continuous distribution of life-span changes in white matter
properties \cite{yeatman2014lifespan}.

\added{We also implemented a union of the PCR Lasso and SGL approaches by first}
\added{projecting each group onto its PC basis. This PCR-SGL approach transformed the feature}
\added{space into a more parsimonious representation of its variance while}
\added{also preserving group structure. But it was unable to efficiently represent}
\added{cross-bundle covariance, which is an advantage of PCR Lasso. As a result, PCR-SGL performed well in ALS classification,}
\added{where the white matter interactions are highly localized, and poorly}
\added{in age regression, where the white matter interactions are global.}

@@

\added{Another limitation of our approach is in its computational execution time,}
\added{which, in our experience, is roughly five times slower than simpler models}
\added{like the lasso. This slow-down is likely to be insignificant for ``production''}
\added{cycles in which researchers report their findings, but may be prohibitive}
\added{for ``development'' cycles in which researchers repeatedly train models while}
\added{adjusting their analysis pipeline. In practice, we circumvent this limitation}
\added{by developing our analysis infrastructure with fast, simple models, and substituting}
\added{slower, more performant models afterward.}

Taken together, \replaced{our}{these} results demonstrate the promise of the
group-regularized regression approach. Even at the scale of dozens of
subjects, the results provided by SGL are both accurate, as well as
interpretable \cite{Murdoch2019-ax}: \replaced{tract profiling}{tractometry} capitalizes on domain
knowledge to engineer meaningful features; SGL scores these features based on
their relative importance; enables a visualization of these feature
importance scores in the anatomical coordinate frame of the bundles (e.g.,
\cref{fig:class-results:coefs3d,fig:age-results:wh-coefs3d,fig:age-results:hbn-coefs3d,fig:age-results:cc-coefs3d})
and provides a means to understand model errors (e.g.,
\cref{fig:class-results:false-positives,fig:class-results:false-negatives}).
Thus, this multivariate analysis approach achieves high cross-validated
accuracy for precision medicine applications of dMRI data and identifies
relevant features of brain anatomy that can further our neuroscientific
understanding of clinical disorders.

Neuroscience has entered an era in which consortium efforts are amassing
large datasets of high-quality dMRI measurements to address a variety of
scientific questions \cite{jernigan2016ping, jernigan2018abcd,
alexander2017open, Miller2016-hw, VanEssen2012}, but the volume and
complexity of these data pose substantial challenges.
\replaced{Tract profiling}{Tractometry} followed by
analysis with SGL provide a promising approach to distill meaningful insights
from the wealth of data measured in these efforts.

@@

\item Diffusion MRI data from the Cambridge Centre for Ageing and
Neuroscience (Cam-CAN) ``CC700'' dataset
\cite{shafto2014cambridge,taylor2017cambridge}, containing data from 640
subjects with ages 18-88. These data were measured on a 3T Siemens TIM Trio
system and written informed consent was obtained from each participant. Voxel
resolution was \num{2x2x2}$\text{mm}^3$ with 30 non-colinear directions
measured for each of $b=1000$ $\text{s} / \text{mm}^2$ and $b=2000$
$\text{s} / \text{mm}^2$. The diffusion weighted images were acquired
with a twice refocused spin-echo sequence and the same preprocessing pipeline
used for the HBN dataset was applied to this data. We will refer to this
dataset as Cam-CAN.

@@

\subsubsection*{\added{SGL with principal components}}
